Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-05-28 thirdPaper/draft/typeConeAssociahedra.tex
corrected proof of Lemma 4.50

\begin{proof}
The almost-split conflation induces an exact sequence
\[
F\susp^{-1}Y\to F\susp^{-1}Z \to FX \to FY\to FZ\to F\susp X.
\]
Because~$X$ is concentrated in non-positive degrees,~$FX=0$.
Let us verify that the map~$F\susp^{-1}Y\to F\susp^{-1}Z$ is surjective.
Let~$\Lambda \xrightarrow{f} \susp^{-1} Z$ be a morphism in~$K^b(\proj \Lambda)$.
Since~$Y\overset{b}{\to} Z$ is right almost-split, and~$Z$ is indecomposable not in~$\add(\susp\Lambda)$, the morphism~$\susp f$ factors through~$b$.
This implies the required surjectivity, and we have a short exact sequence
\[
 0 \to FX \xrightarrow{Fa} FY\xrightarrow{Fb} FZ\to 0.
\]
Because~$\MOD\Lambda$ and~$K^b(\proj \Lambda)$ are Krull--Schmidt, it is immediate to check that~$Fb$ (resp.~$Fa$) inherits from~$b$ (resp.~$a$) the property of being right (resp. left) almost-split.
\end{proof}

@@

\begin{lemma}
Let~$\Lambda$ be an Artin algebra, and let~$Y\infl E\defl X$ be a conflation in~$K^{[-1,0]}(\proj\Lambda)$. Then:
\begin{enumerate}
 \item If~$\Hom{K^b}(Y,\susp E)=0$, we have~$\Hom{\Lambda}(FX,\tau FY)=0$.
 \item If moreover~$Z\notin\add(\susp\Lambda)$ and the conflation is almost-split, then~$\Hom{\Lambda}(FX,\tau^2 FX)=0$.
\end{enumerate}
\end{lemma}

@@

\begin{lemma}\label{lemm::initial meshes}
Let~$\Lambda$ be an Artin algebra whose indecomposable injective modules are~$\tau$-rigid, and let~$P$ be an indecomposable projective~$\Lambda$-module.
%which is a brick.
% Let~$I$ be indecomposable injective with socle~$P/\operatorname{rad}P$, and~$\alpha$ be the composition~$P\to\operatorname{soc} I \to I$.
% Complete the morphism~$\alpha$ to a triangle~$P\to Y\to E\to \susp P$ in~$K^{[-1,0]}(\proj\Lambda)$, where~$Y$ is a minimal projective presentation of~$I$.
% Then~$Y\infl E\defl \susp P$ is an Auslander--Reiten conflation in~$K^{[-1,0]}(\proj\Lambda)$, which is also a mutation conflation.
Then the Auslander--Reiten conflation~$Y\infl E\defl \susp P$ in~$K^{[-1,0]}(\proj\Lambda)$ is a mutation conflation if~$FY$ is a brick.
\end{lemma}
\pg{This seems false for~$\Lambda = k[x]/(x^2)$: AR sequence~$\Lambda\to (\Lambda\to \Lambda) \to \Sigma\Lambda$.  Probably need brick algebra.}

@@

% We first show that the conflation is almost-split.
% 
% blahblahblah

@@

(0) Since~$P$ is concentrated in only one degree, it is rigid in~~$K^{[-1,0]}(\proj\Lambda)$ and so is~$\susp P$.
Moreover,~$FY=I$ is~$\tau$-rigid, hence~$Y$ is rigid in~$K^{[-1,0]}(\proj\Lambda)$ by~\cite[Lemma 3.4]{AdachiIyamaReiten}.

@@

(3) $\Hom{K^b}(E,\susp E)=0$: There is an exact sequence
\[
 \Hom{K^b}(\susp Y,\susp E) \to \Hom{K^b}(\susp P, \susp E) \to \Hom{K^b}(E,\susp E) \to \Hom{K^b}(Y,\susp E)
\]
in which the last term is zero by (2).
The first morphism is isomorphic to~$\Hom{K^b}(Y,E) \to \Hom{K^b}(P,E)$, which is surjective by the previous lemma.

(4) $\Hom{K^b}(E,\susp Y) = 0$: Consider the exact sequence
\[
 \Hom{K^b}(E,E) \to \Hom{K^b}(E,\susp P) \to \Hom{K^b}(E,\susp Y) \to \Hom{K^b}(E,\susp E).
\]
Its last term vanishes by (3) and its first morphism is surjective because~$E\to \susp P$ is right almost-split.

(5) $\Hom{K^b}(\susp P,\susp E)= 0$:
Let~$P\xrightarrow{f} E$ be any morphism.
Since~$P$ is rigid,~$f$ factors through the inflation~$\beta$ in the conflation~$Y\infl E\defl \susp P$, and there is some morphism~$g$ such that~$f=\beta g$.
By the previous lemma, or by (4), the morphism~$g$ factors through~$P \xrightarrow{\alpha} Y$: there is some~$h:Y\to Y$ such that~$g= h \alpha$.\pg{Fixable if~$\Lambda$ is brick? Will think more...}

We then have~$f = \beta g = \beta (g-\alpha) = \beta (h-1) \alpha$.
Since~$Y$ is endo-local, the morphism~$h$ or~$h-1$ is an automorphism.

{\bf Small gap:} There is an~$h'$ such that~$h \alpha = \alpha h'$ and we deduce that~$f=\beta h \alpha = \beta \alpha h' = 0$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{theo::brick-algebra-condition}]
 The theorem follows from the slightly more general statements of Lemma~\ref{lemm::then-AR-seq-are-mutation} and Lemma~\ref{lemm::initial meshes}.
\end{proof}

>>>>

\begin{proof}
The almost-split conflation induces an exact sequence
\[
F\susp^{-1}Y\to F\susp^{-1}Z \to FX \to FY\to FZ\to F\susp X.
\]
Because~$X$ is concentrated in non-positive degrees,~$F\susp X=0$.
Let us verify that the map~$F\susp^{-1}Y\to F\susp^{-1}Z$ is surjective.
Let~$\Lambda \xrightarrow{f} \susp^{-1} Z$ be a morphism in~$K^b(\proj \Lambda)$.
Since~$Y\overset{b}{\to} Z$ is right almost-split, and~$Z$ is indecomposable not in~$\add(\susp\Lambda)$, the morphism~$\susp f$ factors through~$b$.
This implies the required surjectivity, and we have a short exact sequence
\[
 0 \to FX \xrightarrow{Fa} FY\xrightarrow{Fb} FZ\to 0.
\]
Because~$\MOD\Lambda$ and~$K^b(\proj \Lambda)$ are Krull--Schmidt, it is immediate to check that~$Fb$ (resp.~$Fa$) inherits from~$b$ (resp.~$a$) the property of being right (resp. left) almost-split.
\end{proof}

@@

\begin{lemma}\label{lemm::if AR are mutations}
Let~$\Lambda$ be an Artin algebra, and let~$Y\infl E\defl X$ be a conflation in~$K^{[-1,0]}(\proj\Lambda)$. Then:
\begin{enumerate}
 \item If~$\Hom{K^b}(Y,\susp E)=0$, we have~$\Hom{\Lambda}(FX,\tau FY)=0$.
 \item If moreover~$X\notin\add(\susp\Lambda)$ and the conflation is almost-split, then~$\Hom{\Lambda}(FX,\tau^2 FX)=0$.
\end{enumerate}
\end{lemma}

@@

\begin{lemma}\label{lemm::initial meshes}
Let~$\Lambda$ be an Artin algebra whose indecomposable injective modules are~$\tau$-rigid, and let~$P$ be an indecomposable projective~$\Lambda$-module.
%which is a brick.
% Let~$I$ be indecomposable injective with socle~$P/\operatorname{rad}P$, and~$\alpha$ be the composition~$P\to\operatorname{soc} I \to I$.
% Complete the morphism~$\alpha$ to a triangle~$P\to Y\to E\to \susp P$ in~$K^{[-1,0]}(\proj\Lambda)$, where~$Y$ is a minimal projective presentation of~$I$.
% Then~$Y\infl E\defl \susp P$ is an Auslander--Reiten conflation in~$K^{[-1,0]}(\proj\Lambda)$, which is also a mutation conflation.
Then the Auslander--Reiten conflation~$Y\infl E\defl \susp P$ in~$K^{[-1,0]}(\proj\Lambda)$ is a mutation conflation if~$FY$ is a brick.
\end{lemma}

@@

@@

(0) Since~$P$ is concentrated in only one degree, it is rigid in~~$K^{[-1,0]}(\proj\Lambda)$ and so is~$\susp P$.
Moreover,~$FY$ is indecomposable and injective.
By assumption, it is~$\tau$-rigid, hence~$Y$ is rigid in~$K^{[-1,0]}(\proj\Lambda)$ by~\cite[Lemma 3.4]{AdachiIyamaReiten}.

@@

(3) $\Hom{K^b}(E,\susp Y) = 0$: Consider the exact sequence
\[
 \Hom{K^b}(\susp Y,\susp Y) \to \Hom{K^b}(\susp P,\susp Y) \to \Hom{K^b}(E,\susp Y) \to \Hom{K^b}(Y,\susp Y).
\]
Its last term vanishes by (0).
Let us prove that its first morphism is surjective.
Composition with~$\alpha$ and~$F\alpha$ yields a commutative square
\[
\xy
%
(-20,7)*+{\Hom{K^b}(Y,Y)}="0";
(20,7)*+{\Hom{K^b}(P,Y)}="2";
(-20,-7)*+{\Hom{\Lambda}(FY,FY)}="10";
(20,-7)*+{\Hom{\Lambda}(FP,FY)}="12";
%
{\ar@{->>} "0";"10"};
{\ar^{\alpha^\ast} "0";"2"};
{\ar@{->>} "2";"12"};
{\ar_{(F\alpha)^\ast} "10";"12"};
%
{\ar@{}|\circlearrowright "0";"12"};
\endxy
\]
where the vertical maps are surjective.
Since~$\Lambda$ is rigid and~$Y$ is two-term, we can apply~\cite[Proposition 6.2]{IyamaYoshino} to the triangulated category~$K^b(\proj\Lambda)$ (or Proposition~\ref{prop::KRextricat} to the extriangulated~$K^{[-1,0]}(\proj\Lambda)$).
As a consequence, the right-most vertical map is bijective.
Because~$FY$ is a brick,~$\End{\Lambda}(FY)$ is a division ring.
Moreover,~$FY$ is the indecomposable injective with simple socle the simple top of~$P$, so that the~$\End{\Lambda}(FY)$-module~$\Hom{\Lambda}(FP,FY)$ is one-dimensional over~$\End{\Lambda}(FY)$.
It follows that the vertical map~$(F\alpha)^\ast$ is either zero or bijective.
But~$\alpha$ is non-zero with domain in~$\add\Lambda$ so that~$F\alpha$ is non-zero ; and~$(F\alpha)^\ast(1_{FY})=F\alpha$ so that~$(F\alpha)^\ast$ is non-zero.
By commutativity, this implies the surjectivity of the top vertical morphism~$\alpha^\ast$.
Since the shift functor is an automorphism, the morphism~$\Hom{K^b}(\susp Y,\susp Y) \to \Hom{K^b}(\susp P,\susp Y)$ is also surjective, and~$\Hom{K^b}(E,\susp Y) = 0$.

(4) $\Hom{K^b}(E,\susp E)=0$: In the exact sequence
\[
 \Hom{K^b}(E,\susp Y) \to \Hom{K^b}(E, \susp E) \to \Hom{K^b}(E,\susp^2 P),
\]
the first term is zero by (3) and the last term vanishes since~$P$ is concentrated in degree zero.

(5) $\Hom{K^b}(\susp P,\susp E)= 0$:
Let~$P\xrightarrow{f} E$ be any morphism.
\[
\xy
%
(7,14)*+{P}="P";
%
(-21,0)*+{P}="0";
(-7,0)*+{Y}="2";
(7,0)*+{E}="4";
(21,0)*+{\susp P}="6";
%
(-21,-14)*+{Y}="Y";
(-7,-14)*+{E}="E";
%
{\ar^f "P";"4"};
{\ar^0 "P";"6"};
{\ar@{-->}_g "P";"2"};
%
{\ar^\alpha "0";"2"};
{\ar^\beta "2";"4"};
{\ar "4";"6"};
%
{\ar_g "0";"Y"};
{\ar@{-->}_h "2";"Y"};
{\ar_\beta "Y";"E"};
{\ar_{\beta h\!} "2";"E"};
{\ar@{-->}^{h'} "4";"E"};
\endxy
\]
Since~$P$ is rigid,~$f$ factors through the inflation~$\beta$ in the conflation~$Y\infl E\defl \susp P$, and there is some morphism~$g$ such that~$f=\beta g$.
By (3), the morphism~$g$ factors through~$P \xrightarrow{\alpha} Y$: there is some~$h:Y\to Y$ such that~$g= h \alpha$.
We then have~$f = \beta g = \beta (g-\alpha) = \beta (h-1) \alpha$.
Since~$Y$ is endo-local, the morphism~$h$ or~$h-1$ is an automorphism, and we may assume that~$h$ is an automorphism.
The morphism~$\beta h$ is not a section (otherwise~$\beta$ would be a section) and~$\beta$ is left-almost split: there is an endomorphism~$h'$ of~$E$ such that~$\beta h = h'\beta$.
Therefore, we have~$f=\beta h \alpha = h'\beta\alpha = 0$.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{theo::brick-algebra-condition}]
 The theorem follows from the slightly more general statements of Lemma~\ref{lemm::then-AR-seq-are-mutation},~\ref{lemm::if AR are mutations} and Lemma~\ref{lemm::initial meshes}.
\end{proof}
